# Reproducible Research: Peer Assessment 1
Author: BellyTheMagnificent  
[Linkedin](https://www.linkedin.com/pub/lee-chun-wai/24/8/8b6/)    
Date: 2014-06-14  

## Loading and preprocessing the data  
Read data from csv file.  
If the file is not available in the current directory, download from github and unzip it.
``` {r, echo=TRUE}
fileName = "activity.csv"
if(!file.exists(fileName)){
        url <- "https://github.com/BellyTheMagnificent/RepData_PeerAssessment1/blob/master/activity.zip"
        zipName <- "activity.zip"
        download.file(url, zipName, method = "curl")
        unzip(zipName)
}

activity <- read.csv(fileName, stringsAsFactors = FALSE)
```  
Convert the date to date data type  
        1.Convert 'date' column to date data type  
        2.Get weekday name from converted date column  
        3.Combine date and interval into datetime  
``` {r}
activity$date = as.Date(activity$date, "%Y-%m-%d")
activity$day = as.character(activity$date, '%a')
activity$datetime= strptime(paste(activity$date, "00:00:00", sep = ' '), "%Y-%m-%d %H:%M:%S",tz="GMT") + activity$interval * 60
```
Data frame after prepoccess 
```{r, echo=FALSE}
str(activity)
```

## What is mean total number of steps taken per day?
First, let look at the histogram of the total number of steps taken each day. The date with 0 steps due to no data were collected on the particular day.
``` {r, fig.height=5}
bplot = barplot(tapply(activity$steps, activity$date, sum, na.rm=T), ylab="Total steps", 
        xlab = "Date",
        main = "Steps taken between 2012/10/01 to 2012/11/30",
        xaxt='n'
        )
axis(1, at = bplot, labels = F)
```
### Mean
```{r}
meanSteps = mean(tapply(activity$steps, activity$date, sum, na.rm=T))
```
The mean of the steps taken each day is `r meanSteps`.

### Median
```{r}
medianSteps = median(tapply(activity$steps, activity$date, sum, na.rm=T))
```
The median is `r medianSteps`.  

## What is the average daily activity pattern?

To view the average daily activity pattern, first have to create a aggregate data set.
```{r}
aggByInterval = aggregate(steps ~ interval, data = activity, mean)
str(aggByInterval)
```
Then we can plot the data using ggplot2. 
```{r, fig.width=10}
library(ggplot2)
ggplot(aggByInterval, aes(interval,steps), title="average number of steps taken") + geom_line() + labs(title = "Average number of steps taken on particular interval")
```

### Which 5-minute interval have highest steps count? Let's find out.
```{r}
aggByInterval[which.max(aggByInterval$steps),]
```

#### The interval 835 has the highest number of average steps per day.  
We can convert this interval to time to check the exact period about it.
```{r}
datetime = as.character(strptime("2010-01-01 00:00:00","%Y-%m-%d %H:%M:%S", tz="GMT")+ 835 * 60,"%I:%M %p")
```
The time for this interval is `r datetime`.  


## Imputing missing values
The data set is not completed. There are some rows do not have steps recorded.
```{r}
summary(activity$steps)
```
Base on the summary above, we found there is `r summary(activity$steps)[7]` records are having NA as the values for steps column.



## Are there differences in activity patterns between weekdays and weekends?
